---
title: "Final project"
author: "Jieun Park"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(readr)
library(igraph)
library(dplyr)
library(ggplot2)
library(purrr)
library(tibble)
library(caret)
library(tidyverse)
```

# Function

```{r}
sim_sir <- function(g,beta,mu,seeds){
  state <- rep(0,vcount(g)) #initial state of the simulation
  state[seeds] <- 1 #infect the seeds
  t <- 0
  table <- data.frame(t=0,inf=seeds)
  while(sum(state==1)>0){
    t <- t + 1
    ## I -> R
    infected <- which(state==1) #get them
    # generate a random value for every infected, if it's < mu, let the node recover.
    state[infected] <- ifelse(runif(length(infected)) < mu,2,1)
    
    ## S -> I
    infected <- which(state==1)
    susceptible <- which(state==0) #get them
    contacts <- as.numeric(unlist(adjacent_vertices(g,infected))) #get the contacts of infected
    contacts <- contacts[contacts %in% susceptible] # get those who are susceptible
    new_infected <- contacts[runif(length(contacts)) < beta] #infect contacts
    if(length(new_infected)>0){
      state[new_infected] <- 1
      table <- rbind(table,data.frame(t,inf=new_infected))
    }
  }
  table
}
```


# Data collection

```{r}
node <- read.csv("nodes.csv") |> select("X..index", "meta")
edge <- read.csv("edges.csv")

graph <- graph_from_data_frame(edge, vertices = node, directed = FALSE)
graph <- simplify(graph)

cluster <- clusters(graph)
giant_component <- which.max(cluster$csize)
node_giant <- V(graph)[cluster$membership == giant_component]

g <- induced_subgraph(graph, node_giant)
g

plot(g, vertex.label = NULL, vertex.size = 3)
```

# 1. Finding epidemic threshold

```{r}
mu <- 0.1
betac <- mu*mean(degree(g))/(mean(degree(g)^2))
betac
```

# 2. Simulating the SIR model

```{r}
seeds <- sample(1:vcount(g),vcount(g)*0.01)

realization <- sim_sir(g,beta = 0.5,mu=0.1,seeds) %>% 
  group_by(t) %>% summarize(ninf=n()) 

realization %>% 
  ggplot(aes(x=t,y=ninf)) + geom_line()
```

## plot SIR with several beta

```{r}
results <- map_dfr(seq(0,0.2,0.01), # beta range
        \(beta){ seeds <- sample(1:vcount(g),vcount(g)*0.01)
        realization <- sim_sir(g,beta,mu,seeds) # make a realization for each beta
        data.frame(beta,ninf=nrow(realization)) # create dataframe row
        })

results %>% ggplot(aes(x=beta,y=ninf))+ geom_point()+
  geom_vline(xintercept = betac,linetype=2) 
```

## 3. Centrality Proxies

```{r}
realization <- sim_sir(g,beta=0.1,mu=0.1,seeds)
```


### degree

```{r}
table_d <- degree(g) %>% enframe(name = "inf",value="degree") %>%
  merge(realization) 
table_d %>% 
  ggplot(aes(x=t,y=degree)) + geom_point() + scale_y_log10()
```


### closeness

```{r}
table_c <- closeness(g) %>% enframe(name = "inf",value="closeness") %>%
  merge(realization) 
table_c %>% 
  ggplot(aes(x=t,y=closeness)) + geom_point() + scale_y_log10()
```

### betweeness

```{r}
table_b <- betweenness(g) %>% enframe(name = "inf",value="betweenness") %>%
  merge(realization) 
table_b %>% 
  ggplot(aes(x=t,y=betweenness)) + geom_point() + scale_y_log10()
```
### page rank

```{r}
table_pr <- page_rank(g)$vector %>% enframe(name = "inf",value="page_rank") %>%
  merge(realization) 
table_pr %>% 
  ggplot(aes(x=t,y=page_rank)) + geom_point() + scale_y_log10()
```

```{r}
table <- merge(table_d,table_c)
table <- merge(table,table_pr)
table <- merge(table,table_b)
cor(table$t,table[,-c(1,2)])
```

```{r}
seeds <- sample(1:vcount(g),vcount(g)*0.01)
realization_random <- sim_sir(g,beta = 0.5,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
seeds <- closeness(g) %>% enframe(name = "inf",value="closeness") %>%
  slice_max(closeness,n=round(vcount(g)*0.01,0))

seeds_inf <- seeds$inf
seeds_inf <- as.integer(seeds_inf)
all(seeds_inf %in% V(g)) 

realization_targeted <- sim_sir(g,beta = 0.5,mu=0.1,seeds_inf) %>%
  group_by(t) %>% summarize(ninf=n())
ggplot() + geom_line(data=realization_random,aes(x=t,y=ninf,col="Random"))+
  geom_line(data=realization_targeted,aes(x=t,y=ninf,col="Targeted"))

```

```{r}
total_infected_random <- sum(realization_random$ninf)
total_infected_targeted <- sum(realization_targeted$ninf)
difference_infected <- total_infected_targeted - total_infected_random

print(paste("Total Infected (Random)", total_infected_random))
print(paste("Total Infected (Targeted)", total_infected_targeted))
print(paste("Difference in Infected", difference_infected))
```

### Vaccinate people not to spread the trust

```{r}
seeds <- sample(1:vcount(g),vcount(g)*0.01)
realization <- sim_sir(g,beta = 0.5,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
vaccinated_random <- sample(1:vcount(g),vcount(g)*0.05)
gp <- delete_vertices(g,vaccinated_random)
seeds <- sample(1:vcount(gp),vcount(gp)*0.01)
realization_random <- sim_sir(gp,beta = 0.5,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
```

### Choose the seeds with largest centrality

```{r}
seeds <- closeness(g) %>% enframe(name = "inf",value="closeness") %>%
  slice_max(closeness,n=round(vcount(g)*0.05,0))
gp <- delete_vertices(g,seeds$inf)
seeds <- sample(1:vcount(gp),vcount(gp)*0.01)
realization_targeted <- sim_sir(gp,beta = 0.5,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
```

```{r}
ggplot() + geom_line(data=realization,aes(x=t,y=ninf,col="No Vaccination")) +
  geom_line(data=realization_random,aes(x=t,y=ninf,col="Vacc. Random"))+
  geom_line(data=realization_targeted,aes(x=t,y=ninf,col="Vacc. Targeted"))
```

```{r}
No_Vaccination <- sum(realization$ninf)
Vacc._Random <- sum(realization_random$ninf)
Vacc._Targeted <- sum(realization_targeted$ninf)

print(paste("No Vaccination", No_Vaccination))
print(paste("Vacc. Random", Vacc._Random))
print(paste("Vacc. Targeted", Vacc._Targeted))

Vaccination_table <- data.frame(vaccination_label = c("Vacc. Random", "Vacc. Targeted", "No Vaccination"),
                                vaccination_value = c(Vacc._Random, Vacc._Targeted, No_Vaccination))
Vaccination_table
```

### Explanation for part c (1%) and part d (5%)

```{r}
print(paste("Total Infected (Random)", total_infected_random))
print(paste("Total Infected (Targeted)", total_infected_targeted))
print(paste("Difference in Infected", difference_infected))

partc <- data.frame("partc_label" = c("Total Infected (Random)", "Total Infected (Targeted)", 
                                          "Difference in Infected"),
                                "partc_value" = c(total_infected_random, total_infected_targeted, difference_infected))

merged_table <- cbind(Vaccination_table, partc)
merged_table
```

With only 1% coverage, random vaccination appears more effective than targeted vaccination, indicating that the targeting strategy might not be optimal or insufficient at this low level. However, at 5% coverage, targeted vaccination drastically reduces the number of infections compared to random vaccination and no vaccination, demonstrating its effectiveness when a sufficient proportion of the network is vaccinated.

This comparison emphasizes that while targeted vaccination can be very effective, it requires an adequate level of coverage to identify and protect the most critical nodes in the network. Random vaccination might provide more immediate benefits at very low coverage levels but does not capitalize on the network's structure as effectively as targeted vaccination at higher coverage levels.

### Sensor spreading

```{r}
all <- table_d %>% group_by(t) %>% summarize(ninf=n())
sensors <- table_d %>% filter(degree>10) %>% group_by(t) %>% summarize(ninf=n())
ggplot() + geom_line(data=all,aes(x=t,y=ninf/sum(ninf),col="All")) + geom_line(data=sensors,aes(x=t,y=ninf/sum(ninf),col="Sensors"))
```

### neighbor

```{r}
control <- sample(1:vcount(g),1000)
get_one_neighbor <- function(id){
  sample(neighbors(g,id),1)
}
sensors <- sapply(control,get_one_neighbor)
```

### cummulative plot

```{r}
control_evo <- table_d %>% filter(inf %in% control) %>% group_by(t) %>% summarize(ninf=n())
sensors_evo <- table_d %>% filter(inf %in% sensors) %>% group_by(t) %>% summarize(ninf=n())
ggplot() + geom_line(data=control_evo,aes(x=t,y=cumsum(ninf/sum(ninf)),col="All")) +
  geom_line(data=sensors_evo,aes(x=t,y=cumsum(ninf/sum(ninf)),col="Sensors")) + labs(y="# infections (cumm)")
```

Control is the random connection and sensors used the friendship paradox which means your friends normally have more friends than you. Therefore, when we have used the sensors proxy then there are more people who got infected than the randomly chosen people. 

## Prediction
### Linear Regression

```{r}
table

# Split the data into training and testing sets
set.seed(123)
train_index <- createDataPartition(table$t, p = 0.8, list = FALSE)
test_data <- table[-train_index, ] |> select(-inf)

# 5-fold cross validation
train_data <- table[train_index, ] |> select(-inf)

control <- trainControl(method = "cv",  # Use k-fold cross-validation
                        number = 5)

# Train a regression model
model <- train(t ~ ., data = train_data, method = "lm", trControl = control)

cv_results <- model$results

```

Since the RMSE is too high, we have used the lasso regularization.

```{r}
x <- as.matrix(train_data[,-1])
y <- train_data$t

control <- trainControl(method = "repeatedcv", number = 10, repeats = 5)

ridge_model <- train(
  x, y,
  method = "glmnet",
  trControl = control,
  tuneGrid = expand.grid(alpha = 0, lambda = 10^seq(-3, 0, length = 100))
)

ridge_model$bestTune

prediciton <- predict(ridge_model, newdata = x)
sqrt(mean((prediciton - y)^2))



# Evaluate the model on the test data
predictions_t <- predict(model, newdata = train_data)
rmse <- sqrt(mean((predictions_t - train_data$t)^2))
print(paste("RMSE on test data:", rmse))

predictions <- predict(model, newdata = test_data)
rmse <- sqrt(mean((predictions - test_data$t)^2))
print(paste("RMSE on test data:", rmse))
```



```{r}

```


