---
title: "social network final project basic code"
author: "Jieun Park"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulating the SIR model 


```{r cars}
sim_sir <- function(g,beta,mu,seeds){
  state <- rep(0,vcount(g)) #initial state of the simulation
  state[seeds] <- 1 #infect the seeds
  t <- 0
  table <- data.frame(t=0,inf=seeds)
  while(sum(state==1)>0){
    t <- t + 1
    ## I -> R
    infected <- which(state==1) #get them
    # generate a random value for every infected, if it's < mu, let the node recover.
    state[infected] <- ifelse(runif(length(infected)) < mu,2,1)
    
    ## S -> I
    infected <- which(state==1)
    susceptible <- which(state==0) #get them
    contacts <- as.numeric(unlist(adjacent_vertices(g,infected))) #get the contacts of infected
    contacts <- contacts[contacts %in% susceptible] # get those who are susceptible
    new_infected <- contacts[runif(length(contacts)) < beta] #infect contacts
    if(length(new_infected)>0){
      state[new_infected] <- 1
      table <- rbind(table,data.frame(t,inf=new_infected))
    }
  }
  table
}
```

## Including Plots

The output of our sim_sir function is a table with the ids of the nodes and when they got infected.

```{r pressure, echo=FALSE}
# Sample a preferential attachnment network
g <- sample_pa(100,directed=F,m=2); 
# Sample an initial infectious seed
seeds <- sample(1:vcount(g),1)
# Generate a simlulation of the SIR model
realization <- sim_sir(g,beta = 0.5,mu=0.1,seeds) 
realization
```


Let’s get a one realization on a bigger preferential-attachment graph (Barabasi) with 1% random seeds.
```{r}
g <- sample_pa(10000,directed=F,m=2)
seeds <- sample(1:vcount(g),vcount(g)*0.01)
realization <- sim_sir(g,beta = 0.5,mu=0.1,seeds) 
```

visualize the infection 

```{r}
realization <- sim_sir(g,beta = 0.5,mu=0.1,seeds) %>% 
  group_by(t) %>% summarize(ninf=n()) 
realization %>% 
  ggplot(aes(x=t,y=ninf)) + geom_line()
```

check epidemic threshold 
```{r}
mu <- 0.1
betac <- mu*mean(degree(g))/(mean(degree(g)^2))
betac
```
 
 Calculating each realization the number of total infections and trying different values of β and fixed μ.
 
```{r}
results <- map_dfr(seq(0,0.2,0.01), # beta range
        \(beta){ seeds <- sample(1:vcount(g),vcount(g)*0.01)
        realization <- sim_sir(g,beta,mu,seeds) # make a realization for each beta
        data.frame(beta,ninf=nrow(realization)) # create dataframe row
        })
results %>% ggplot(aes(x=beta,y=ninf))+ geom_point()+
  geom_vline(xintercept = betac,linetype=2)
```
 
Simulating the SIR model on a Erdos-Renyi graph with the same average degree

```{r}
er <- sample_gnm(n=vcount(g),m=ecount(g),directed = F)
```

calculating beta (threshold)

```{r}
betac_er <- mu*mean(degree(er))/(mean(degree(er)^2))
betac_er
```

Simulating

```{r}
mu <- 0.1
results_er <- map_dfr(seq(0,0.2,0.01),
        \(beta){seeds <- sample(1:vcount(er),vcount(er)*0.01)
          realization <- sim_sir(er,beta,mu,seeds)
          data.frame(beta,ninf=nrow(realization))
        })
ggplot() + geom_line(data=results,aes(x=beta,y=ninf,col="PA"))+
  geom_line(data=results_er,aes(x=beta,y=ninf,col="ER"))
```

### Which nodes get the infection earlier?
Simulating spreading above the threshold 
```{r}
realization <- sim_sir(g,beta=0.1,mu=0.1,seeds)
```

Adding the degree of each node and see the relationship between degree and time to infection:

```{r}
table_d <- degree(g) %>% enframe(name = "inf",value="degree") %>%
  merge(realization) 
table_d %>% 
  ggplot(aes(x=t,y=degree)) + geom_point() + scale_y_log10()
```

#Other types of centrality 

closeness

```{r}
table_c <- closeness(g) %>% enframe(name = "inf",value="closeness") %>%
  merge(realization) 
table_c %>% 
  ggplot(aes(x=t,y=closeness)) + geom_point() + scale_y_log10()
```

betweenness

```{r}
table_b <- betweenness(g) %>% enframe(name = "inf",value="betweenness") %>%
  merge(realization) 
table_b %>% 
  ggplot(aes(x=t,y=betweenness)) + geom_point() + scale_y_log10()
```

page rank 

```{r}
table_pr <- page_rank(g)$vector %>% enframe(name = "inf",value="page_rank") %>%
  merge(realization) 
table_pr %>% 
  ggplot(aes(x=t,y=page_rank)) + geom_point() + scale_y_log10()
```

Which predicts better the time to infection? 
```{r}
table <- merge(table_d,table_c)
table <- merge(table,table_pr)
table <- merge(table,table_b)
cor(table$t,table[,-c(1,2)])
```

##Immunize/seed networks to prevent/accelerate spreading
What will happen if we select our seeds according to centrality?

```{r}
seeds <- sample(1:vcount(g),vcount(g)*0.01)
realization_random <- sim_sir(g,beta = 0.5,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
seeds <- closeness(g) %>% enframe(name = "inf",value="closeness") %>%
  slice_max(closeness,n=vcount(g)*0.01)
realization_targeted <- sim_sir(g,beta = 0.5,mu=0.1,seeds$inf) %>%
  group_by(t) %>% summarize(ninf=n())
ggplot() + geom_line(data=realization_random,aes(x=t,y=ninf,col="Random"))+
  geom_line(data=realization_targeted,aes(x=t,y=ninf,col="Targeted"))
```

Vaccination 

20% of the nodes in the network randomly

```{r}
seeds <- sample(1:vcount(g),vcount(g)*0.01)
realization <- sim_sir(g,beta = 0.5,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
vaccinated_random <- sample(1:vcount(g),vcount(g)*0.2)
gp <- delete_vertices(g,vaccinated_random)
seeds <- sample(1:vcount(gp),vcount(gp)*0.01)
realization_random <- sim_sir(gp,beta = 0.5,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
```

Choosing the seeds with the largest centrality 
```{r}
seeds <- closeness(g) %>% enframe(name = "inf",value="closeness") %>%
  slice_max(closeness,n=vcount(g)*0.2)
gp <- delete_vertices(g,seeds$inf)
seeds <- sample(1:vcount(gp),vcount(gp)*0.01)
realization_targeted <- sim_sir(gp,beta = 0.5,mu=0.1,seeds) %>%
  group_by(t) %>% summarize(ninf=n())
```

Visualizing the evolution together with the realization with no vaccination we got earlier.

```{r}
ggplot() + geom_line(data=realization,aes(x=t,y=ninf,col="No Vaccination")) +
  geom_line(data=realization_random,aes(x=t,y=ninf,col="Vacc. Random"))+
  geom_line(data=realization_targeted,aes(x=t,y=ninf,col="Vacc. Targeted"))
```

##Sensors of the spreading

Selecting a group of nodes that signal when the spreading is happening
```{r}
all <- table_d %>% group_by(t) %>% summarize(ninf=n())
sensors <- table_d %>% filter(degree>10) %>% group_by(t) %>% summarize(ninf=n())
ggplot() + geom_line(data=all,aes(x=t,y=ninf/sum(ninf),col="All")) + geom_line(data=sensors,aes(x=t,y=ninf/sum(ninf),col="Sensors"))
```

Picking a random set of nodes in the network (our control group) and use one random neighbor of them as the sensor group
```{r}
control <- sample(1:vcount(g),1000)
get_one_neighbor <- function(id){
  sample(neighbors(g,id),1)
}
sensors <- sapply(control,get_one_neighbor)
```

Visualization when they get infected (cummulative)

```{r}
control_evo <- table_d %>% filter(inf %in% control) %>% group_by(t) %>% summarize(ninf=n())
sensors_evo <- table_d %>% filter(inf %in% sensors) %>% group_by(t) %>% summarize(ninf=n())
ggplot() + geom_line(data=control_evo,aes(x=t,y=cumsum(ninf/sum(ninf)),col="All")) +
  geom_line(data=sensors_evo,aes(x=t,y=cumsum(ninf/sum(ninf)),col="Sensors")) + labs(y="# infections (cumm)")
```

